---
title: "Overview of Datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The {avocado} package consists of three different datasets that summarize the weekly sales of Hass Avocados at different regional levels.

- `hass_usa`: weekly contiguous US avocado sales at the country level
- `hass_region`: weekly contiguous US avocado sales at the region level
- `hass`: weekly contiguous US avocado sales at the city/sub-region level

The unit sales of avocados is actually measured in terms of weight - not individual avocados. Furthermore, the average price in the dataset is not adjusted for inflation.

### PLU

The product/price lookup code ([PLU](https://en.wikipedia.org/wiki/Price_look-up_code)) uniquely identifies a product (mainly produce). The [Hass Avocado Board](https://loveonetoday.com/how-to/identify-hass-avocados/) focuses on six different PLUs:

* 4046: non-organic small/medium Hass Avocados (~3-5 oz)
* 4225: non-organic large Hass Avocados (~8-10 oz)
* 4770: non-organic extra large Hass Avocados (~10-15 oz)
* 94046: organic small/medium Hass Avocados (~3-5 oz)
* 94225: organic large Hass Avocados (~8-10 oz)
* 94770: organic extra large Hass Avocados

### Bags vs PLU

Another distinction that the HAB makes is between bags versus bulk. Bulk typically means avocados of a similar kind tracked in bulk by weight and are easily distinguishable with their PLU codes.  For instance, a PLU of 4046 means all avocados sold in bulk that are of a certain size (see PLU section). On the other hand, bags indicate a pre-packaged container consisting of a variable number of avocados of mixed PLU type. For instance, a package of six avocados may consist of 2 PLU 4046, 3 PLU 4770 and 1 PLU 4225. In other words, bagged sales are unable to account for individual PLU sales.

### Bulk GTIN

Bulk GTIN is a new 'variable' that was introduced in late 2020 and refers to avocados sold that are not intended for point-of-sale. Examples include businesses that purchase avocados as 'raw materials' for making packaged guacamole.

### Region vs. Location

The `hass_region` and `hass` datasets contain a shared variable called `region` and the `hass` dataset has a variable called `location`. Regions are defined by the Hass Avocado Board and locations are selected cities or sub-regions that are part of the overall region. The totals found for all locations within a Region will not equal the total found for the specific Region due to the aforementioned point. For convenience, here is a breakdown of the Regions and Locations:

* California
* West
  - Washington
  - Oregon
  - Idaho
  - Nevada
  - Montana
  - Utah
  - Arizona
  - Wyoming
  - Colorado
  - New Mexico
* Plains
  - North Dakota
  - South Dakota
  - Nebraska
  - Kansas
  - Minnesota
  - Iowa
  - Missouri
* South Central
  - Texas
  - Oklahoma
  - Arkansas
  - Louisiana
* Southeast
  - Mississippi
  - Alabama
  - Georgia
  - South Carolina
  - Florida
* Midsouth
  - Kentucky
  - Tennessee
  - North Carolina
  - West Virginia
  - Virginia
  - Maryland
  - Delaware
* Great Lakes
  - Wisconsin
  - Illinois
  - Michigan
  - Indiana
  - Ohio
* Northeast
  - Pennsylvania
  - New York
  - Vermont
  - New Hampshire
  - Massachusetts
  - Connecticut
  - Rhode Island
  - New Jersey
  - Maine

## Datasets

### hass_usa

The `hass_usa` dataset focuses on weekly Hass Avocado sales at the country (i.e., contiguous US) level and consists of the following fields:

- `week_ending`: The date of the last day of the week in YYYY-MM-DD format
- `avg_price_nonorg`: The average selling price in US$ for non-organic Hass avocados. Not adjusted for inflation
- `total_nonorg_units`: This is the total unit sales (in US pounds) of non-organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu4046` + `plu4225` + `plu4770` + `total_nonorg_bag_units` + `bulk_nonorg_gtin`
- `total_org_units`: This is the total unit sales (in US pounds) of organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu94046` + `plu94225` + `plu94770` + `total_org_bag_units` + `bulk_norg_gtin`
- `plu4046`: The amount of PLU 4046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4225`: The amount of PLU 4225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4770`: The amount of PLU 4770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_nonorg`-bag: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `avg_price_org`: The average selling price in US$ for organic Hass avocados. Not adjusted for inflation
- `plu94046`: The amount of PLU 94046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94225`: The amount of PLU 94225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94770`: The amount of PLU 94770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `bulk_nonorg_gtin`: The amount of non-organic avocados sold as bulk and not intended for point-of-sales
- `bulk_org_gtin`: The amount of organic avocados sold as bulk and not intended for point-of-sales

```{r}
library(avocado)
data('hass_usa')
dplyr::glimpse(hass_usa)
```

### haas_region

The `hass_region` dataset focuses on weekly US Hass Avocado sales at the region level and consist of the following fields:

- `region`: Specific region within the US as defined by the Hass Avocado Board
- `week_ending`: The date of the last day of the week in YYYY-MM-DD format
- `avg_price_nonorg`: The average selling price in US$ for non-organic Hass avocados. Not adjusted for inflation
- `total_nonorg_units`: This is the total unit sales (in US pounds) of non-organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu4046` + `plu4225` + `plu4770` + `total_nonorg_bag_units` + `bulk_nonorg_gtin`
- `total_org_units`: This is the total unit sales (in US pounds) of organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu94046` + `plu94225` + `plu94770` + `total_org_bag_units` + `bulk_norg_gtin`
- `plu4046`: The amount of PLU 4046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4225`: The amount of PLU 4225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4770`: The amount of PLU 4770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_nonorg`-bag: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `avg_price_org`: The average selling price in US$ for organic Hass avocados. Not adjusted for inflation
- `plu94046`: The amount of PLU 94046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94225`: The amount of PLU 94225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94770`: The amount of PLU 94770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `bulk_nonorg_gtin`: The amount of non-organic avocados sold as bulk and not intended for point-of-sales
- `bulk_org_gtin`: The amount of organic avocados sold as bulk and not intended for point-of-sales

```{r}
library(avocado)
data('hass_region')
dplyr::glimpse(hass_region)
```


### hass

The `hass` dataset summarizes weekly Hass Avocado sales within the contiguous US based on city or sub-region. These areas are defined by the HAB and make up portions of the `region` field in the `haas_region` dataset. The fields are:

- `location`: Specific location within a region of the US. These locations enable a further deep-dive into avocado sales within a region
- `region`: Specific region within the US as defined by the Hass Avocado Board
- `week_ending`: The date of the last day of the week in YYYY-MM-DD format
- `avg_price_nonorg`: The average selling price in US$ for non-organic Hass avocados. Not adjusted for inflation
- `total_nonorg_units`: This is the total unit sales (in US pounds) of non-organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu4046` + `plu4225` + `plu4770` + `total_nonorg_bag_units` + `bulk_nonorg_gtin`
- `total_org_units`: This is the total unit sales (in US pounds) of organic avocados sold as bulk, not bulk, and bags. There is a discrepancy between this value and summing up the individual PLU columns and the individual bag size columns (e.g., large, small, etc.). The values in this column can be derived with summing the columns: `plu94046` + `plu94225` + `plu94770` + `total_org_bag_units` + `bulk_norg_gtin`
- `plu4046`: The amount of PLU 4046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4225`: The amount of PLU 4225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu4770`: The amount of PLU 4770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_nonorg_bag`: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_nonorg`-bag: The amount of non-organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `avg_price_org`: The average selling price in US$ for organic Hass avocados. Not adjusted for inflation
- `plu94046`: The amount of PLU 94046 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94225`: The amount of PLU 94225 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `plu94770`: The amount of PLU 94770 Hass avocados sold by weight in US pounds. This does not include avocados sold in pre-packaged quantities. See vignette for more information about Hass PLUs
- `small_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in small pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `large_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `xlarge_org_bag`: The amount of organic Hass avocados (they can be a mix of PLUs) sold in extra-large pre-packaged containers/bags in US pounds. This does not include avocados sold individually. See vignette for more information about pre-packaged avocados
- `bulk_nonorg_gtin`: The amount of non-organic avocados sold as bulk and not intended for point-of-sales
- `bulk_org_gtin`: The amount of organic avocados sold as bulk and not intended for point-of-sales

```{r}
library(avocado)
data('hass')
dplyr::glimpse(hass)
```


## Discrepancy

You may notice that there are some discrepancies when comparing the values in the total columns (e.g., `total_nonorg_bag_units`) versus the component-like columns (e.g., `xlarge_nonorg_bag`). This discrepancy stems from the underlying data as recorded by the Hass Avocado Board. They obtain data from another organization and there may be discrepancies in how the data are collected.

For instance, the column `total_org_bag_units` may seem to be the total of three columns: `small_org_bag`, `large_org_bag`, and `xlarge_org_bag`. However, the sum of these 3 columns will often be less. On their website, Hass Avocado seems to use the total columns as often as possible without going into details of the totals.

## Source

The raw data are available on [Hass Avocado Board's](https://hassavocadoboard.com/category-data/) website (registration required).
